---
title: "Barents Hg river flux"
output: html_document
date: "2022-10-21"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Calculate River Hg Flux
- Rivers in Norway Målselva, Alta, Tana, and Pasvik
- Rivers in Russia: Northern Dvina and possibly Tuloma

We have monthly Hg samples with monthly water flow. 

# Calculation methods
According to Achmed et al., (2018) doi: 10.1007/s10661-017-6444-y.there are different ways of calculating river flux depending on the sampling frequencey and variable of interest. 

Downside of our data (at least the Russian) has the downside of water flow not available at the time of sampling for Hg determination. Because of this I choose the Global Mean (GM) method. The method is considered more imprecise and inaccurate.  
```{r Load pacages}
Packages <- c("tidyr","plyr", "dplyr", "ggplot2")
lapply(Packages, library, character.only = TRUE)
```
The following files are needed:
- Monthly flow from Norwegian Rivers
- Monthly Hg measure in Norwegian Rivers
- Montly flow with Hg measured in Russian Rivers
```{r Load Data}
#Norwegian data
RM19 <- read.table("221021_MonthlyFlow_Rivermonitoring_2019.txt", sep = "\t", header = TRUE)
RM20 <- read.table("221021_MonthlyFlow_Rivermonitoring_2020.txt", sep = "\t", header = TRUE)
RM21 <- read.table("221021_MonthlyFlow_Rivermonitoring_2021.txt", sep = "\t", header = TRUE)
RMHg <- read.table("221021_RM_THg data212019.txt", sep = "\t", header = TRUE)
#Need to rename station Målselv
RMHg$Station.name <- revalue(RMHg$Station.name, c("M?selva" = "Malselv"))

#Russian data
Rus <- read.table("221021_Russia.txt", sep = "\t", header = TRUE)
```

```{r Data Pre-processing}
No_flow <- rbind(RM19, RM20, RM21)

#Calculate from flow per second to per month
No_flow$DisMean_m3 <- No_flow$DisMean_m3s * 2628288*1000
No_flow$Variable <- "Flow_m3"
colnames(No_flow)[which(names(No_flow) == "DisMean_m3")] <- "Value"

RMHg$Year=gsub(" mean.*","",RMHg$Year)
colnames(RMHg)[which(names(RMHg) == "THg")] <- "Value"
RMHg$Variable <- "THg"

#Rus from wide to long
Rus$DisMean_m3 <- Rus$DisMean_m3s * 2628288*1000
colnames(Rus)[which(names(Rus) == "DisMean_m3")] <- "Flow_m3"
Rus_long <- gather(Rus, Variable, Value, c(THg,Flow_m3), factor_key=TRUE)

AllNor <- rbind(No_flow[,c(3:5, 7:8)], RMHg[,c(1, 3, 5:7)])
AllNor2 <- subset(AllNor, Year == "2019"|Year =="2020")
AllRiv <- rbind(Rus_long[,c(1, 4:5, 7:8)], AllNor2)
#Subset rivers
```

Find annual water discharge to see correlation to Hg
```{r Annual water discharge}
#summarise Value for variable == Flow_m3, by Year and Station.name

bajabaj <- AllRiv %>% 
    filter(Station.name == "Altaelva"| Station.name == "Pasvikelva" | Station.name == "Tanaelva" | Station.name == "Malselv"|Station.name == "Northern Dvina"|Station.name == "Tuloma") %>%
  group_by(Station.name, Year, Variable) %>% 
  summarise(sumr = sum(Value))

#filter(Variable == "Flow") %>%
```


The global mean (GB) method (Huang et al. 2012) calculated annual river load as a product of the total annual flow and average contamninant concentrations for the period of the record, as expressed by the following equation (Shih et al. 1994):

L = mQt(∑(Ci/n)),

L = annual river load (M/T)
m = conversion factor to convert the calculated values into a specific unit
Ci = contaminant concentration (M/L3) measured at the ith day
Qt = total annual flow (L3/T)
n = number of samples

Considering to use this method on a monthly basis and then to summarise. 
monthly flow * contaminant concentration

OBS!! make sure the units are correct

- FLOW: for each month, multiply by 2,628,288 seconds per month and summarise per site per year to get Q of m3/y
```{r Calculations}

#Norwegian rivers
summary(AllRiv)
AllRiv$Value <- as.numeric(AllRiv$Value)
typeof(AllRiv$Value)

#Calculate annual sum of water flow
df1 <- AllRiv %>%
  group_by(Station.name, Year) %>%
  summarize(sumFlow = sum(Value[Variable=="Flow_m3"]))

df2 <- AllRiv %>%
  group_by(Station.name, Year) %>%
  summarize(sumHg = mean(Value[Variable=="THg"], na.rm = T))

total <- merge(df1, df2,by=c("Station.name", "Year"))

class(total$Station.name)

total2 <- total %>%
  filter(Station.name == "Altaelva"| Station.name == "Pasvikelva" | Station.name == "Tanaelva" | Station.name == "Malselv") %>%
  group_by(Station.name, Year) %>%
  mutate(FLOW = (sumFlow*sumHg)*10^-15)

#Russian RIvers
summary(Rus_long)

#Calculate annual sum of water flow
df1x <- Rus_long %>%
  group_by(Station.name, Year) %>%
  summarize(sumFlow = sum(Value[Variable=="Flow_m3"]))

df2x <- Rus_long %>%
  group_by(Station.name, Year) %>%
  summarize(sumHg = mean(Value[Variable=="THg"], na.rm = T))

totalx <- merge(df1x, df2x, by=c("Station.name", "Year"))

total2x <- totalx %>%
  group_by(Station.name, Year) %>%
  mutate(FLOW = (sumFlow*sumHg)*10^-15)



```
## To make a nice plot
- plot boxplot with NO and RU rivers THg per month with flow as line
```{r Organise the data}
#Norwegian flow data
#No_flow 
#RMHg
#fjern kolonne 7
NOR_flow2 = select(No_flow, -7)
colnames(NOR_flow2)[which(names(NOR_flow2) == "DisMean_m3s")] <- "Value"

AllNOR2 <- rbind(NOR_flow2[,c(3:7)], RMHg[,c(1, 3, 5:7)])

#Rus
Rus_long2 <- gather(Rus, Variable, Value, c(THg,DisMean_m3s), factor_key=TRUE)

AllNOR2$Variable2 <- "Norway"
Rus_long2$Variable2 <- "Russia"

AllBAR <- rbind(AllNOR2, Rus_long2[,c(1, 4:5, 7:9)])
AllBAR$Variable <- gsub("Flow_m3", "DisMean_m3s", AllBAR$Variable)
```


```{r make the plot}
AllBAR2 <- filter(AllBAR, Station.name == "Altaelva"| Station.name == "Pasvikelva" | Station.name == "Tanaelva" | Station.name == "Malselv"| Station.name == "Northern Dvina" | Station.name == "Tuloma") 

AllBAR2$Value <- as.numeric(AllBAR2$Value)
ggplot(AllBAR2, aes(x=as.numeric(Month), y=Value, group=as.numeric(Month), colour=Station.name))+
  geom_boxplot()+
  geom_jitter(width=0.15)+
  facet_grid(Variable~Variable2, scales = "free")
 
       
summary(AllBAR2)


ggplot(df, aes(x=group, y=values, fill=group)) +
  geom_boxplot() +
  stat_summary(fun=mean, geom='point', shape=20)

```

```{r export for supporting tables}
write.csv(AllBAR2, file = "Flow and Hg.csv")

```

```{r Make plot dual y-axis}

library(ggplot2)
library(dplyr)
library(patchwork) # To display 2 charts together
library(hrbrthemes)

#Preparation
AllBAR3 <- spread(AllBAR2, Variable, as.numeric(Value))
#AllBAR3$type <- as.factor(ifelse(AllBAR3$Station.name == "Northern Dvina", "4",
#                          ifelse(AllBAR3$Station.name == "Malselv", "1", 
#                          ifelse(AllBAR3$Station.name == "Altaelva", "1", 
#                          ifelse(AllBAR3$Station.name == "Tanaelva", "2",
#                          ifelse(AllBAR3$Station.name == "Pasvikelva", "2",
#                          ifelse(AllBAR3$Station.name == "Tuloma", "2", "")))))))
#Average flow data per year and type?
#Make plot with 2019 and 2020 data, facet by station name
#AllBAR3x <- filter(AllBAR3, !Year == c("2019")) 
#AllBAR3x <- filter(AllBAR3x, !Year == c("2021")) 

#ggplot(AllBAR3x, aes(x=Month, colour=Station.name)) +
#  geom_bar( aes(y=THg), stat="identity")+
#  geom_line(aes(y=DisMean_m3s/300), size=0.8) +
#  facet_wrap(Station.name~., scales="free")

#Averages
#ggplot(AllBAR6, aes(x=Month, colour=Station.name)) +
#  geom_line(aes(y=mean_Dis/300), size=0.8) +
#  geom_point( aes(y=mean_THg, group=Month))+
#  facet_wrap(Station.name~., scales="free")

#DENNE ER BESTE SÅ LANGT. BRUKER MEDIAN AV TO ÅR
#make median of flow and THg 
AllBARy <- na.omit(AllBAR3) %>% 
  filter(!Year == "2021") %>% 
  group_by(Month, Station.name)%>%
  summarise(mean_Dis=median(DisMean_m3s))

AllBAR7 <- merge(AllBARy, AllBAR3, by=c("Station.name","Month"))
AllBARx <- na.omit(AllBAR3) %>% 
  filter(!Year == "2021") %>% 
  group_by(Month, Station.name)%>%
  summarise(mean_THg=median(THg))


AllBAR6 <- merge(AllBARy, AllBARx, by=c("Station.name","Month"))
Sys.setlocale("LC_CTYPE", locale="Norwegian") #To allow for å in Måløselv
AllBAR6$Station.name <- revalue(AllBAR6$Station.name, c("Malselv" = "Målselv", "Tanaelva"="Tana",  "Altaelva" = "Alta", "Pasvikelva"="Pasvik"))


#Function for å definere antall siffer i akse
fmt_dcimals <- function(decimals=1){
  function(x) format(x,nsmall = decimals,scientific = FALSE)
}



AllBAR6 %>%
  arrange(Station.name) %>%
  mutate(Station.name = factor(Station.name, levels=c("Målselv", "Alta", "Tana", "Pasvik", "Tuloma", "Northern Dvina"))) %>%
  filter(!Station.name == c("Tuloma")) %>% 
  filter(!Station.name == c("Northern Dvina")) %>% 
  ggplot(aes(x=Month, colour=Station.name)) +
  geom_bar(aes(y=mean_THg), fill="lightgrey",stat="identity")+
  geom_line(aes(y=mean_Dis/300), size=1.3) +
  facet_wrap(Station.name~.)+
  scale_x_discrete(name ="Month", 
                    limits=c("1","2","3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))+
  scale_y_continuous(name = "THg (ng/L)", labels = fmt_dcimals(1), breaks=c(0, 1, 2, 3),
    # Add a second axis and specify its features
    sec.axis = sec_axis( trans=~.*300, name="Water Discharge (m3/s)"))+
  theme_bw()+
  theme(legend.position = "none",
       axis.text.y = element_text(size= 16, colour="black"),
          axis.title.x = element_blank(),
          axis.title.y = element_text(size=18, margin=margin(0,10,0,0)),
          axis.text.x = element_text(size = 16, colour="black"),
          legend.title = element_blank(),
          legend.text=element_text(size=16),
          legend.spacing.x = unit(0.2, 'cm'),
          strip.text = element_text(size=16),
       plot.margin=unit(c(1,3,1,1), "cm"),
       axis.title.y.right = element_text(vjust=4))
  ylim(0, 3)+
  scale_y_continuous(labels = fmt_dcimals(1))



#Northern DVina alone due to high discharge
AllBAR6 %>%
  filter(Station.name == c("Northern Dvina")) %>% 
  ggplot(aes(x=Month, colour="CD9600")) +
  geom_bar(aes(y=mean_THg),colour="#CD9600", fill="lightgrey", stat="identity")+
  geom_line(aes(y=mean_Dis/1300),colour="#CD9600", size=1.7) +
  scale_x_discrete(name ="Month", 
                    limits=c("1","2","3", "4", "5", "6", "7", "8", "9", "10", "11", "12"))+
  scale_y_continuous(name = "THg (ng/L)",labels = fmt_dcimals(1),
    # Add a second axis and specify its features
    sec.axis = sec_axis( trans=~.*1300, name="Water Discharge (m3/s)")
  )+
  theme_bw()+
  facet_wrap(Station.name~.)+
  theme(legend.position = "none",
            axis.text.y = element_text(size= 16, colour="black"),
          axis.title.x = element_blank(),
          axis.title.y = element_text(size=18, margin=margin(0,10,0,0)),
          axis.text.x = element_text(size = 16, colour="black"),
          legend.title = element_blank(),
          legend.text=element_text(size=16),
          legend.spacing.x = unit(0.2, 'cm'),
          strip.text = element_text(size=16),
       plot.margin=unit(c(1,3,1,1), "cm"),
       axis.title.y.right = element_text(vjust=4))

  ylim(0, 3)
  

```

